The importance of measurement uncertainty in ecological management

### 2018-01-30 manuscripts/manuscript.Rmd
Revisions from milad

---
title: "The importance of measurement uncertainty in ecological management"
author:
  - name: "Milad Memarzadeh"
    affiliation: a
  - name: "Carl Boettiger"
    affiliation: a, 1
address:
  - code: a
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"

abstract: |
  We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.  Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to intuition and the precautionary principle. Through the use of cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.  In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value through adaptive learning.

@@

bibliography: citations.bib
#journal: "The American Naturalist"
date: "`r Sys.Date()`"
csl: the-american-naturalist.csl
output: rticles::elsevier_article
layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column
header-includes: |
        \journal{The American Naturalist}

@@

```{r optional_plot_fonts, message=FALSE}
library(ggplot2)
library(Cairo)
library(extrafont)
library(hrbrthemes)
extrafont::loadfonts()
knitr::opts_chunk$set()
hrbrthemes::import_roboto_condensed() 
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
palette <- c("#D9661F", "#3B7EA1",  "#6C3302", 
             "#FDB515", "#00B0DA",  "#CFDD45")
colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: high prior",
                      "biomass",
                      "catch"))
scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
```

@@

We seek to determine the sequence of actions $h_t$ for $t \in [1, ... \infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.  We will denote the discount rate $\gamma$. For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax. Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,

@@

Given an estimated recruitment model along with this measurement of the stock size, the manager must set a harvest quota $h_t$ for that year in order to maximize their expected long term (net) present value:

$$\max_{ \lbrace h_t \rbrace} \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$

@@

A second subtly is in the nature of the decision problem itself: the optimization problem must always consider the set ${h_t}$ all future actions simultaneously, not merely the next action.  This is a game of chess, where it can be optimal to make costly short term actions (such as closing the fishery) if they pay off in the long run. This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a single harvest quota $H_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.  Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.

@@

$$x_{t+1} = x_t r  \left(1 - \frac{x_t}{K}\right) - h_t \varepsilon_t$$

@@

However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.  Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g. @].  As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.

@@

Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].  Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].  Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.  Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-fishing Limit, (OFL). Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment. To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.  To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$. To distinguish this approach from MSY, we will refer to this approach as a TAC policy.  This more closely represents a heuristic [e.g. @Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.  Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.  Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.

@@

So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of of paradoxes.  Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible. Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.  (This intuition must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.  While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.  Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above. Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarize here.

@@

## Clark's Paradox: $M < D$

@@

- Transition function (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.  In our context, this relationship is given by the Graham-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $\mathcal{N_T}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.

- Observation function: $O(x_t,y_t,a_t)$ the probability of observing state $y$ given a system in state $x$.  In principle, the action chosen can influence the precision of the observation. In our case, we simply assume normally distributed errors around the true state, $y \sim \mathcal{N_T}(x, \sigma_m)$, truncated at zero to exclude negative population sizes.

- Utility function: $U(x_t,a_t)$, the utility received at time $t$ for taking action $a_t$, given that the system is in state $x_t$.  For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t - a_t, 0)$, indicating that realized harvest cannot be negative.  This choice ensures that in the case of no uncertainty ($\sigma_g = \sigma_m = 0$), the optimal solution matches that expected under a simple MSY calculation. More realistic utility functions may include diminishing returns with increasing harvest (supply and demand effects), and the cost of fishing, both of which act to suppress large harvests. By focusing on a simple utility we can be sure that our comparison to MSY is driven by the treatment of uncertainty rather than merely differing economic assumptions.

The optimization problem is to select the action $a_t$ that will maximize the net present utility over all time. Future utility may be discounted by a factor $\gamma$, so that a value  $V$ in $t$ year is valued at $\gamma^t V$ today.  Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  While the solution method for MDP problems, stochastic dynamic programming (SDP) is well known and readily implemented in ecological problems [e.g @Mangel1988; @Marescot2013], the POMDP problem cannot be efficiently solved using SDP. While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intelligence [. @Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.

@@

```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$. Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{{sims}}. The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error. The more cautious TAC is not economically optimal but largely unimpacted by increasing error, while POMDP attains consistently high economic yield despite the increasing uncertainty. \\label{econ}"}
sims %>%
  select(time, value, rep, method, sigma_m, sigma_g) %>%
  filter(sigma_g == "0.15", method != "MSY") %>%
  group_by(rep, method, sigma_m, sigma_g) %>%
  summarise(npv = sum(value)) %>%  
  group_by(method, sigma_m, sigma_g) %>%
  summarise(net_value = mean(npv), se = sd(npv) / mean(npv)) %>%
  ungroup() %>%

@@

This difference between underestimating and overestimating accounts for the poor performance of the CE solution under large measurement uncertainty, where it over-harvests whenever measurements are too large.  Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where chance local extinction becomes possible.  It is precisely this asymmetry: that too much over-harvesting leads to an irreversible state of extinction, while too much under-harvesting is always reversible (modulo some lost revenue) that lay behind Clark's original intuition that there was something fishy about Reed's result that $S = D$: that uncertainty required no extra caution.  Constant escapement is particularly susceptible to over-harvesting starting from stock sizes much higher than $B_{MSY}$ since its bang-bang optimal solution attempts to bring stocks back to target level as fast as possible. CE is not so vulnerable to collapse from small stock sizes, since it shuts down all harvests once estimates fall below $B_{MSY}$. Stock collapse under TAC can also be driven by such a string of heads, but is unlikely at high stock sizes since harvest never exceeds $H_{MSY}$.  In both cases, some measurement uncertainty interacts with inherent stochasticity, which provides a continued source of variation to sizes above $B_{MSY}$, where CE strategy is most vulnerable, and below $B_{MSY}$, where TAC is most vulnerable.

@@

So far we have focused on simulation and an examination of the relative policies under different prior information.  Figure \ref{historical} compares the harvest level that would have been recommended by each of the strategies we have considered here against the historically observed catch recorded for two commercially fished stocks: Argentine Hake and European Plaice.  Historical estimates of biomass and catch are taken from the R.A. Myers Legacy Stock Assessment database [@RAM]. Posterior distributions for parameters for the Graham-Schaefer model are estimated with uncertainty through Markov Chain Monte Carlo using `nimble` [@nimble] on the historical data, illustrating how this process might be done in more complex models as well (details and code in Supplement, Appendix B).  $B_{MSY} and corresponding TAC and CE policies are calculated based on posterior mean estimates of the model parameters, along with POMDP solution assuming a 10% measurement error rate.  Though measurement error could be estimated directly from the raw data, this would not reflect the true measurement uncertainty arising from the stock assessment process.  Each strategy is then compared to historical observations to determine a recommended harvest.  As we have seen, the TAC and CE harvest policies are uniquely determined by the observed stock size relative to $B_{MSY}$, but the POMDP policy must be re-calculated each time step to reflect both the prior observation and prior action.  Note that since we cannot replay history under these alternative policies, so the prior action taken is always the historical catch for that year and not the catch recommended by the POMDP policy.

@@

First, it may seem obvious that POMDP would out-preform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.  Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accommodate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.  If stochasticity doesn't matter, it is easy to believe that measurement error would be minor at best.  In fact, we find that once we account for measurement error, the effects of both stochasticity and measurement have a significant influence over the optimal policy.  It is similarly tempting to dismiss the better performance of POMDP against a more rule-of-thumb[^1] based strategy like the TAC scenario considered here (fishing at 80% MSY), but once again this was not a forgone conclusion.  While a better economic return might reasonably be expected of the optimization approach, it was not obvious that POMDP would also be at times more cautious than the rule of thumb, able to eek out both higher economic returns by harvesting more when it was safe to do so, and higher ecological recovery by harvesting even less than TAC when necessary.  These results highlight the difficulty of finding effective heuristic rules and underscore the power and utility of more sophisticated strategies such as POMDP which can make the best use of all available information.

@@

Several limitations that have been studied in fully observed (MDP) optimal
decision problems, such as parameter uncertainty [e.g. @Ludwig1982], model 
uncertainty [@Williams2001; @Boettiger2015] and adaptive management [e.g. @Walters1976]
remain largely open challenges for partially observed systems. Future work 
could extend this analysis to more complex models, such as those with age structure,
as [@Holden2015] does for the fully observed case. Another limiting assumption 
common to MDPs and POMDPs is that of stationary dynamics: that the population 
dynamics equation itself is not changing over time. In reality, forces such as
climate change and other forms of environmental variations violate this
assumption  [@Britten2017]. Direct approaches such as @Fackler2014's
adaptation of Mixed Observability MDP [@Ong2010] do not scale to the number
of states and actions considered here. A value of information (VOI) analysis
for POMDP,such as @Johnson2015 shows for MDP models, could identify when it
is worthwhile to actively reduce measurement error.

>>>>

---
title: "The importance of measurement uncertainty in ecological management"

author:
  - name: "Authors removed during peer review"

#  - name: "Milad Memarzadeh"
#    affiliation: a
#  - name: "Carl Boettiger"
#    affiliation: a, 1
#address:
#  - code: a
#    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"

abstract: |

  We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.  Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to intuition and the precautionary principle. Through the use of cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.  In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value through adaptive learning.

@@

bibliography: citations.bib
#journal: "The American Naturalist"
date: "`r Sys.Date()`"
csl: the-american-naturalist.csl
output: rticles::elsevier_article
layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column
header-includes: |
        \journal{The American Naturalist}
        \usepackage{lineno}
        \linenumbers
        \modulolinenumbers[3]
        \usepackage{setspace}
        \doublespacing

@@

```{r optional_plot_fonts, message=FALSE}
library(ggplot2)
library(Cairo)
library(extrafont)
library(hrbrthemes)
extrafont::loadfonts()
knitr::opts_chunk$set()
#hrbrthemes::import_roboto_condensed() 
#ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
ggplot2::theme_set(hrbrthemes::theme_ipsum())
palette <- c("#D9661F", "#3B7EA1",  "#6C3302", 
             "#FDB515", "#00B0DA",  "#CFDD45")
colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: high prior",
                      "biomass",
                      "catch"))
scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
```

@@

We seek to determine the sequence of actions $h_t$ for $t \in [1, ... \infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.  We will denote the discount rate $\gamma$. For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax. Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size subject to some measurement uncertainty $\xi_t$

@@

For simplicity we will assume measurement uncertainty is normally distributed with standard deviation $\sigma_m$,  $\xi \sim \mathcal{N}(1, \sigma_m)$.  Given an estimated recruitment model along with this measurement of the stock size, the manager must choose the set of a harvest quotas $\lbrace h_t \rbrace$ for $t \in [0, 1, \infty]$ that maximizes expected long term (net) present value:

$$\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$

@@

A second subtly is in the nature of the decision problem itself: note that the manager in not myopic, thinking only about the next year, but rather considers the sequence of all future actions ${h_t}$.  While the manager will revisit this sequence each year in light of new observations, it is this ability of the decision problem to think ahead that allows it to tolerate short-term costs (e.g. reduced harvests in the current season) for larger future payoffs.  In this, ecological management is a game of chess, always thinking several moves ahead. This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a fixed mortality $F_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.  Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.

@@

$$x_{t+1} = f(x_t, h_t) = \varepsilon_t x_t r  \left(1 - \frac{x_t}{K}\right) - h_t $$

@@

However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.  Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g. @Dichmont2016].  As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.

@@

Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].  Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].  Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.  Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-fishing Limit, (OFL). Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment. To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.  To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under the MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$. To distinguish this approach from MSY, we will refer to this approach as a TAC policy.  This more closely represents a heuristic [e.g. @Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.  Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.  Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.

@@

So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of paradoxes.  Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible. Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.  (This intuition must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.  While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.  Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above. Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarize here.

@@

## Clark's Paradox

@@

- Transition function (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.  In our context, this relationship is given by the Gordon-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.

- Observation function: $O(x_t,y_t,a_{t-1})$ the probability of observing state $y_t$ given a system in state $x_t$.  In principle, the action chosen can influence the precision of the observation. In our case, we simply assume normally distributed errors around the true state, $y_t \sim \mathcal{N}(x_t, \sigma_m)$, truncated at zero to exclude negative population sizes.

- Utility function: $U(x_t,a_t)$, the utility received at time $t$ for taking action $a_t$, given that the system is in state $x_t$.  For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t, a_t)$, indicating that realized harvest cannot be negative.  This choice ensures that in the case of no uncertainty ($\sigma_g = \sigma_m = 0$), the optimal solution matches that expected under a simple MSY calculation. More realistic utility functions may include diminishing returns with increasing harvest (supply and demand effects), and the cost of fishing, both of which act to suppress large harvests. By focusing on a simple utility we can be sure that our comparison to MSY is driven by the treatment of uncertainty rather than merely differing economic assumptions.

The optimization problem is to select the action $a_t$ that will maximize the net present utility over all time. Future utility may be discounted by a factor $\gamma$, so that a value  $V$ in $t$ year is valued at $\gamma^t V$ today.  Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  While the solution method for MDP problems, stochastic dynamic programming (SDP) is well known and readily implemented in ecological problems [e.g @Mangel1988; @Marescot2013], the POMDP problem cannot be efficiently solved using SDP. While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intelligence [@Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.

@@

```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$. Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{sims}. The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error. The more cautious TAC is not economically optimal but largely unimpacted by increasing error, while POMDP attains consistently high economic yield despite the increasing uncertainty. \\label{econ}"}
sims %>%
  select(time, value, rep, method, sigma_m, sigma_g) %>%
  filter(sigma_g == "0.15", method != "MSY") %>%
  group_by(rep, method, sigma_m, sigma_g) %>%
  summarise(npv = sum(value)) %>%  
  group_by(method, sigma_m, sigma_g) %>%
  summarise(net_value = mean(npv), se = sd(npv) / mean(npv)) %>%
  ungroup() %>%

@@

This difference between underestimating and overestimating accounts for the poor performance of the CE solution under large measurement uncertainty, where it over-harvests whenever measurements are too large.  Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where the chance of local extinction becomes possible.  It is precisely this asymmetry: that too much over-harvesting leads to an irreversible state of extinction, while too much under-harvesting is always reversible (modulo some lost revenue) that lay behind Clark's original intuition that there was something fishy about Reed's result that $S = D$: that uncertainty required no extra caution.  Constant escapement is particularly susceptible to over-harvesting starting from stock sizes much higher than $B_{MSY}$ since its bang-bang optimal solution attempts to bring stocks back to target level as fast as possible. CE is not so vulnerable to collapse from small stock sizes, since it shuts down all harvests once estimates fall below $B_{MSY}$. Stock collapse under TAC can also be driven by such a string of heads, but is unlikely at high stock sizes since harvest never exceeds $H_{MSY}$.  In both cases, some measurement uncertainty interacts with inherent stochasticity, which provides a continued source of variation to sizes above $B_{MSY}$, where CE strategy is most vulnerable, and below $B_{MSY}$, where TAC is most vulnerable.

@@

So far we have focused on simulation and an examination of the relative policies under different prior information.  Figure \ref{historical} compares the harvest level that would have been recommended by each of the strategies we have considered here against the historically observed catch recorded for two commercially fished stocks: Argentine Hake and European Plaice.  Historical estimates of biomass and catch are taken from the R.A. Myers Legacy Stock Assessment database [@RAM]. Posterior distributions for parameters for the Graham-Schaefer model are estimated with uncertainty through Markov Chain Monte Carlo using `nimble` [@nimble] on the historical data, illustrating how this process might be done in more complex models as well (details and code in Supplement, Appendix B).  $B_{MSY}$ and corresponding TAC and CE policies are calculated based on posterior mean estimates of the model parameters, along with POMDP solution assuming a 10% measurement error rate.  Though measurement error could be estimated directly from the raw data, this would not reflect the true measurement uncertainty arising from the stock assessment process.  Each strategy is then compared to historical observations to determine a recommended harvest.  As we have seen, the TAC and CE harvest policies are uniquely determined by the observed stock size relative to $B_{MSY}$, but the POMDP policy must be re-calculated each time step to reflect both the prior observation and prior action.

@@

First, it may seem obvious that POMDP would out-preform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.  Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accommodate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.  If stochasticity doesn't matter, it is easy to believe that measurement error would be minor at best.  In fact, we find that once we account for measurement error, the effects of both stochasticity and measurement have a significant influence over the optimal policy.  It is similarly tempting to dismiss the better performance of POMDP against a more rule-of-thumb[^1] based strategy like the TAC scenario considered here (fishing at 80% MSY), but once again this was not a forgone conclusion.  While a better economic return might reasonably be expected of the optimization approach, it was not obvious that POMDP would also be at times more cautious than the rule of thumb, able to deliver both higher economic returns by harvesting more when it was safe to do so, and higher ecological recovery by harvesting even less than TAC when necessary.  These results highlight the difficulty of finding effective heuristic rules and underscore the power and utility of more sophisticated strategies such as POMDP which can make the best use of all available information.

@@

Several limitations that have been studied in fully observed (MDP) optimal
decision problems, such as parameter uncertainty [e.g. @Ludwig1982], model 
uncertainty [@Williams2001; @Boettiger2015] and adaptive management [e.g. @Walters1976]
remain largely open challenges for partially observed systems. Future work 
could extend this analysis to more complex models, such as those with age structure,
as [@Holden2015] does for the fully observed case. Another limiting assumption 
common to MDPs and POMDPs is that of stationary dynamics: that the population 
dynamics equation itself is not changing over time. In reality, forces such as
climate change and other forms of environmental variations violate this
assumption  [@Britten2017]. Direct approaches such as @Fackler2014's
adaptation of Mixed Observability MDP [@Ong2010] do not scale to the number
of states and actions considered here. A value of information (VOI) analysis
for POMDP [e.g. @Johnson2015; @Memarzadeh2016b], could identify when it
is worthwhile to actively reduce measurement error.
